Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is a merged and final version of two conference papers, \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}, with a unified and streamlined presentation and expanded background materials.}

### 2020-08-24 journal paper/content/sec-intro.tex
move comment to model seciton

We consider two extremes: a simple Bayesian model with full Bayesian rationality and no performance signals, and a simple frequentist model with reputation scores and no prior knowledge or beliefs. For the theoretical results, the ``no-performance-signals" assumption makes agents' behavior independent of a particular realization of the prior. Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior.\gaedit{\footnote{While our model assumes that agents have correct beliefs about the performance of the algorithms, only the beliefs themselves matter and so our results hold as long as all agents have the same initial beliefs, even if they are not aligned with the true prior.}}  Such summarization is essential for formulating lucid and general analytic results, let alone proving them. It is unclear how to incorporate performance signals in a theoretically tractable model. For the numerical results, the \ExptsModel accounts for competition in a more direct way \gaedit{by positing a more realistic model of consumer behavior that does not require them to have direct information on the algorithms deployed or the bandit problem faced by the firm. It further allows us to separate reputation vs. data advantage and makes our model amenable to numerical simulations.} Indeed, simulating the intricate interplay of learning dynamics and Bayesian priors appears computationally intractable.

>>>>

We consider two extremes: a simple Bayesian model with full Bayesian rationality and no performance signals, and a simple frequentist model with reputation scores and no prior knowledge or beliefs. For the theoretical results, the ``no-performance-signals" assumption makes agents' behavior independent of a particular realization of the prior. Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior.  Such summarization is essential for formulating lucid and general analytic results, let alone proving them. It is unclear how to incorporate performance signals in a theoretically tractable model. For the numerical results, the \ExptsModel accounts for competition in a more direct way \gaedit{by positing a more realistic model of consumer behavior that does not require them to have direct information on the algorithms deployed or the bandit problem faced by the firm. It further allows us to separate reputation vs. data advantage and makes our model amenable to numerical simulations.} Indeed, simulating the intricate interplay of learning dynamics and Bayesian priors appears computationally intractable.

### 2020-08-24 journal paper/content/sec-model.tex
Added a label.
Changed the placement of a figure.
Moved "background" to Appendix A.
move comment to model seciton

\xhdr{Agents' response.} Each agent $t$ forms a reward estimate $\est_i(t)\in [0,1]$ for each principal $i$, and chooses the principal based on these reward estimates. Specifically, principal $1$ is chosen with probability
\begin{align}
p_t = \respF\rbr{ \est_1(t) - \est_2(t) },
\end{align}
where $\respF:[-1,1]\to [0,1]$ is the \emph{response function}, same for all agents and known to everyone.

@@

\begin{figure}
\begin{center}
  \begin{tikzpicture}[scale=2.0]
    \draw[->] (-1.1,0) -- (1.1,0) node[right]
    {$\Delta_t := \est_1(t) - \est_2(t)$};
    \draw[->] (0,-0.1) -- (0,1.1) node[above]
        {$p_t = \text{prob. of choosing principal 1}$};
    \draw[scale=1.0,domain=-1:0,smooth,variable=\q,blue, line width=0.50mm] plot ({\q},{0});
    \draw[scale=1.0,domain=0:1,smooth,variable=\q,blue,line width=0.50mm] plot ({\q},{1});
    \draw[scale=1.0,domain=-1:0,smooth,variable=\y,red]  plot ({\y},{0.1});
    \draw[scale=1.0,domain=0:1,smooth,variable=\y,red]  plot ({\y},{0.9});
    \draw[scale=1.0,domain=-1:1,smooth,variable=\y,cyan, line width=0.45mm, dash pattern=on 3pt off 2pt]  plot ({\y},{1/(1 + 1/(9^\y))});
    % \node[above, blue] at (0.5, 0.5) {\footnotesize $2 q (1 - q)$};
     \node[left] at (0, 0.5) {\footnotesize $1/2$};
     \node[left] at (0, 1) {\footnotesize $1$};
     \node[below left] at (0, 0) {\footnotesize $0$};
     \node[below ] at (1, 0) {\footnotesize $1$};
     \node[below ] at (-1, 0) {\footnotesize $-1$};
  \end{tikzpicture}
\end{center}
\caption{The three models for $\respF$: \HardMax is thick blue, \HardMaxRandom is slim red, and \SoftMaxRandom is the dashed curve.}
\label{fig:response-functions}
\end{figure}

@@

\begin{itemize}
\item In the \emph{\TheoryModel}, the mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from a common Bayesian prior $\priorMu$. Agents know the principals' algorithms. Their reward estimates are defined as posterior mean rewards:
        $\est_i(t) = \PMR_i(t) := \E\sbr{ r_t\mid i_t = i }$
    for each agent $t$ and principal $i$.

>>>>

\xhdr{Agents' response.} Each agent $t$ forms a reward estimate $\est_i(t)\in [0,1]$ for each principal $i$, and chooses the principal based on these reward estimates.gaedit{\footnote{While our model assumes that agents have correct beliefs about the performance of the algorithms, only the beliefs themselves matter and so our results hold as long as all agents have the same initial beliefs, even if they are not aligned with the true prior.}} Specifically, principal $1$ is chosen with probability
\begin{align}\label{eq:model-f}
p_t = \respF\rbr{ \est_1(t) - \est_2(t) },
\end{align}
where $\respF:[-1,1]\to [0,1]$ is the \emph{response function}, same for all agents and known to everyone.

@@

\begin{figure}[t]
\begin{center}
  \begin{tikzpicture}[scale=2.0]
    \draw[->] (-1.1,0) -- (1.1,0) node[right]
    {$\Delta_t := \est_1(t) - \est_2(t)$};
    \draw[->] (0,-0.1) -- (0,1.1) node[above]
        {$p_t = \text{prob. of choosing principal 1}$};
    \draw[scale=1.0,domain=-1:0,smooth,variable=\q,blue, line width=0.50mm] plot ({\q},{0});
    \draw[scale=1.0,domain=0:1,smooth,variable=\q,blue,line width=0.50mm] plot ({\q},{1});
    \draw[scale=1.0,domain=-1:0,smooth,variable=\y,red]  plot ({\y},{0.1});
    \draw[scale=1.0,domain=0:1,smooth,variable=\y,red]  plot ({\y},{0.9});
    \draw[scale=1.0,domain=-1:1,smooth,variable=\y,cyan, line width=0.45mm, dash pattern=on 3pt off 2pt]  plot ({\y},{1/(1 + 1/(9^\y))});
    % \node[above, blue] at (0.5, 0.5) {\footnotesize $2 q (1 - q)$};
     \node[left] at (0, 0.5) {\footnotesize $1/2$};
     \node[left] at (0, 1) {\footnotesize $1$};
     \node[below left] at (0, 0) {\footnotesize $0$};
     \node[below ] at (1, 0) {\footnotesize $1$};
     \node[below ] at (-1, 0) {\footnotesize $-1$};
  \end{tikzpicture}
\end{center}
\caption{The three models for $\respF$: \HardMax is thick blue, \HardMaxRandom is slim red, and \SoftMaxRandom is the dashed curve.}
\label{fig:response-functions}
\end{figure}

@@

\begin{itemize}
\item In the \emph{\TheoryModel}, the mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from a common Bayesian prior $\priorMu$. Agents know the principals' algorithms.\gaedit{\footnote{While our model assumes that agents have correct beliefs about the performance of the algorithms, only the beliefs themselves matter and so our results hold as long as all agents have the same initial beliefs, even if they are not aligned with the true prior.}} Their reward estimates are defined as posterior mean rewards:
        $\est_i(t) = \PMR_i(t) := \E\sbr{ r_t\mid i_t = i }$
    for each agent $t$ and principal $i$.
